# Introduction {#intro}

## Motivation

This document is written for users of the Manifesto Project Data. 
This introduction aims at three goals. It should

1. provide an overview of the data maintained and distributed by the Manifesto Project. 
2. give hands-on advice on how to use the data based on the current methodological discussion. 
3. illustrate how the data by the Manifesto Project was used in the past to answer a large variety of research questions. 

Currently, many information on the project as well as the scholarly debate is fragmented and distributed across very different types of documents and publications. This book attempts to bring this distributed information together and thereby give a low-level introduction to scholars that are completely unfamiliar with the Manifesto Project. Scholars and students that are already familiar with the data or who have already used it in the past can certainly skip various chapters and read the book more selectively.  


examples:

learn how to use the manifesto project dataset to create graphs and tables

```{r, echo = FALSE, warning = FALSE, fig.cap = 'Positions of German parties at the federal elections in 2009.'}
mpds %>%
  filter(countryname == "Germany", date == 200909) %>%
  arrange(party) %>%
  mutate(
    partyabbrev = ifelse(party==41113,"Greens",partyabbrev)
  ) %>%
  ggplot(aes(x = rile, y = date)) +
    xlab("Left-Right Position") +
    ylab("") +
    scale_colour_manual("Party", values = c("CDU/CSU" = "black", "SPD" = "red", "Greens" = "green3", "LINKE" = "deeppink3", "FDP" = "gold")) +
    geom_vline(xintercept = 0, colour = "black", size=0.7) +
    geom_hline(yintercept = 200909.0, colour = "black", size=0.5, alpha=0.5) +
    geom_point(size=3, aes(color = partyabbrev)) +
    #geom_text(aes(label=partyabbrev, color = partyabbrev,  vjust = -0.9), position = position_dodge(10))+
    geom_label(aes(label=partyabbrev, color = partyabbrev,  vjust = -0.9), position = position_dodge(7))+
    theme(axis.ticks = element_blank(), axis.text.y = element_blank(), legend.position = "none")+
    coord_fixed(ratio = 10)

```

and how to analyze the Manifesto Corpus.

## Brief History of the Project {#history}

The Manifesto Project is one of the longest-running projects in the field of party research.



### Manifesto Research Group (MRG)
The Manifesto Research Group (MRG) was founded in 1979 by David Robertson and Ian Budge. Robertson's comparative study on electoral manifestos in the UK and party platforms in the US was the theoretical and empirical starting point [@robertson1976].

MRG was a group of researchers with the common interest to analyze parties' electoral programs in their own countries. At the beginning, the group was mostly formed by scholars from OECD countries. The first published output of the research group was published in 1987 [@budge1987]. The book was mostly structured in a one-chapter/one-country format indicating the most

The data collection was largely decentralized. Although there was agreement among the members of the group on a common coding scheme, the collection and coding was conducted by each researcher individually. 

### Comparative Manifestos Project (CMP)
In 1989 the work started by MRG was continued as the Comparative Manifestos Project under the directorship of Hans-Dieter Klingemann at the WZB Berlin Social Science Center. The CMP pursued the collection and coding efforts of the MRG, but with a stronger centralization of the data collection. The until then collected data was pooled and published on CD-ROM along with the book Mapping Policy Preferences I [@budge2001-1]. With the Manifesto Coding Instructions, Mapping Policy Preferences also contained a detailed coding handbook that made the coding procedure more transparent. The CMP extended the scope of the project and included many central and easten European countries in the analysis. A second version of the pooled dataset was published in 2006 as a companion to Mapping Policy Preferences II [@klingemann2006]. 

### Manifesto Research on Political Representation (MARPOR)

In 2009, the Manifesto Research on Political Representation (MARPOR) a DFG-funded infrastructure project continued the work and data collection of the CMP. 
With the large funding of the DFG, the infrastructure workflows were completely redeveloped and completely digitalized so that MARPOR now provides access to all raw texts. 
MARPOR publishes a new version of the [Manifesto Project Main Dataset](#maindataset) twice a year. Moreover, since 2015 MARPOR publishes the machine-readable texts and codings in the form of the [Manifesto Corpus](#Corpus). 

## Structure of the Book

* The second chapter discusses the [methodological approach](#methodology) from the case sampling  of countries, parties and documents over the coding procedure including the category scheme to the scaling and measurement of parties' left-right positions.
* The third chapter on [scholarly usage](#usage) presents an overview of applications of the manifesto summaries of existing
* The following chapter explains [how to work with the Manifesto Project Dataset](#maindataset). It includes explanations on how to access the dataset, the dataset structure, and how to use the R package manifestoR and the Stata-addon manifestata to apply various scaling procedures.
* The chapter on [Manifesto Corpus](#corpus) lays out the fundamentals of the Manifesto Corpus. Access, structure and exemplary applications are presented.
* The sixth chapter explains the functionality of the [Manifesto API](#api): the programming interface of the Manifesto Project that allows standardized and structured access to all data provided by the Manifesto Project. 
* The [final chapter](#conclusion) provides further readings on the Manifesto Project as well as  contact information.





<!--

You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].

-->
