---
title: "Using the Manifesto Project Dataset to scale party preferences"
always_allow_html: yes
output: 
  md_document:
    variant: markdown_strict
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(tibble)
library(manifestoR)
library(DT)
library(pander)

mp_setapikey("manifesto_apikey.txt")

corpus_version <- "2017-2"
mpds_version <- "2017b"

mpds <- mp_maindataset(version = mpds_version)


```

# Introduction

This tutorial explains the structure and variables of the Manifesto Project Dataset. The goal of this tutorial is that users learn how to use the Manifesto Project Dataset to scale and visualize differences in party preferences over time and between parties.


## Example Dataset


Each row in the dataset represents one electoral program. **The variables party and date uniquely identify every row in the dataset.**


```{r, message=FALSE, warning=FALSE, paged.print=TRUE}

cmp_cats <- paste("per",v4_categories(),sep="")

mpds %>% filter(countryname=="United States" & date == 201211) %>%
  DT::datatable(caption="Extract Main Dataset", rownames=FALSE, options = list(scrollX = TRUE)) %>%
  formatStyle(c('country','countryname','oecdmember','eumember'), color='white',backgroundColor = 'blue', fontWeight = 'bold') %>%
  formatStyle(c('edate','date'), color='white', backgroundColor = 'green', fontWeight = 'bold') %>%
  formatStyle(c('party','partyname','partyabbrev','parfam'), color='white', backgroundColor = 'red', fontWeight = 'bold') %>%
  formatStyle(c('coderid','manual','coderyear','testresult','testeditsim'), color='white', backgroundColor = 'orange', fontWeight = 'bold') %>%
  formatStyle(c('pervote','voteest','presvote','absseat','totseats'), color='black', backgroundColor = 'yellow', fontWeight = 'bold') 
```



## Country and election variables



## Content analytical categories





## Using specific categories

for now we just use the standard three-digit content analytical categories. See the tutorial on subcategories for how to use the other per-categories.


```{r}
library(ggplot2)

mpds %>% filter(countryname=="United States" & date == 201211) %>% select(countryname,date,partyname,per107) %>% pander()

ggplot(mpds %>% filter(countryname=="United States" & date > 198000) %>% select(countryname,date,partyname,per501)) +
  geom_line(aes(x=date,y=per501,z=partyname))
```

## Ready-made indices: Left-right positions

rile

other measures

## More measures: Nicheness


nicheness measure



party family


clarity, 

## create your own scale

average across parties







## save as dta and open in Stata or Excel





## bootstrap

some documents are very short

good way to ...
