---
title: "First steps with manifestoR"
author: "Jirka Lewandowski <jirka.lewandowski@wzb.eu> & Nicolas Merz <nicolas.merz@wzb.eu>"
date: "17 May 2018"
output:
  rmdformats::html_clean:
    highlight: kate
---

*This tutorial is largely based on the [manifestoR vignette](https://cran.r-project.org/web/packages/manifestoR/vignettes/manifestoRworkflow.pdf). *


# Installing and loading manifestoR

In order to make use of the manifestoR package, you need to have a working installation of R installed on your computer. R is an open-source statistic and programming software and can be [downloaded for free](https://cran.r-project.org/mirrors.html). We recommend the use of [RStudio](https://www.rstudio.com/products/rstudio/) as an integrated development environment, but the tutorial should also work with just R installed on your computer. To install the manifestoR-package, type the following commands into the console window:

```{r message=FALSE, warning=FALSE, paged.print=FALSE, eval=FALSE}
install.packages("manifestoR")
```

To make use of manifestoR, you need to load the package using the `library`-function:

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(manifestoR)
```

You need to load the package again whenever you restart your R session, but you do not need to install it again. Although you can type all commands in the R console and use R interactively, we recommend to write all code in an .R file to make your work more easily reproducible. 

# Connecting to the Manifesto Project Database API

To access the Manifesto Project Data or data in the Manifesto Corpus with manifestoR, an account for the Manifesto Project webpage with an API key is required.
If you do not yet have an account, you can create one at 
[https://manifesto-project.wzb.eu/signup]().
If you have an account, you can create and download the API key on your profile page.

For every R session using manifestoR and connecting to the Manifesto API,
you need to set the API key in your work environment.
This can be done by passing either a key or the name of a file containing the
key to manifestoR's `mp_setapikey()` function (see documentation `?mp_setapikey`
for details).
Thus, your R script using manifestoR usually will start like this:

```{r}
library(manifestoR)
mp_setapikey("manifesto_apikey.txt")
```
```{r eval=TRUE, include=FALSE}
library(devtools)
#document("../")
#load_all("../")
#mp_setapikey(key.file = "../tests/manifesto_apikey.txt")
```

This R code presumes that you have stored and downloaded the API key in a file
named `manifesto_apikey.txt` in your current R working directory.

Note that it is a security risk to store the API key file or a script containing
the key in public repositories such as github.

All the following commands will only work if you have set your apikey as described above.


# Downloading the Manifesto Project Dataset

Let's first learn how to download the [Manifesto Project Dataset](https://manifesto-project.wzb.eu/tutorials/main-dataset) with manifestoR. You can download the Manifesto Project Dataset (MPDS) with the function `mp_maindataset()`. 

```{r, message=FALSE, warning=FALSE}
mpds <- mp_maindataset()
names(mpds)
```

The dataset is returned as a data.frame (to be precise as a [tibble](https://tibble.tidyverse.org/) - the tidyverse version of a data.frame), therefore the `names` function returns the variable names of the data frame. By default the most recent version of the dataset is returned, but you can also access older versions. To get a list of all versions of the main dataset, type:

```{r}
mp_coreversions()
```

To query a specific version of the main dataset, use the dataset.id listed in the output of coreversions. For example, to get the Manifesto Project Dataset version 2015a, type:

```{r, eval=FALSE}
mp_maindataset(version = "MPDS2015a")
```

Even if you want to use the current version of the dataset, it is good practice to specifically query this version to ensure that you always get the same dataset even if you later run your script again. 

You can get the Manifesto Project South America Dataset using the function `mp_southamerica_dataset()` that works analogously to `mp_maindataset()`.

# Downloading documents from the Manifesto Corpus


## Check the availability of documents 

Before downloading documents, the function `mp_availability` let's you check which documents are available in the Manifesto Corpus. The following command summarizes the availability of documents. If not assigned to an object, it prints a summary report of the Manifesto Corpus. The argument `TRUE` here indicates that it checks whether there is a document available for each case of the Manifesto Project Dataset. 

```{r}
mp_availability(TRUE)
```

Instead of indicating `TRUE`, you can also indicate a logical expression using variables from the Manifesto Project Dataset which serves as a reference of cases. The following command checks for the availability of documents for all Belgium manifestos covered by the Manifesto Project Dataset. 
 
```{r}
available_docs <- mp_availability(countryname=="Belgium")
available_docs
names(available_docs)
```

available_docs is a data.frame where can easily be filtered, eg for a specific language. To check for the availability of Flemish (here labelled as "dutch") documents for Belgium elections since 2010 covered in the dataset you could for example do the following:

```{r}
belgium_2010 <- mp_availability(countryname=="Belgium" & date > 201000)
filter(belgium_2010, language=="dutch")
```

To get all English-language documents that come along with annotations, you would do the following

```{r}
english_annotated <- mp_availability(TRUE) %>% filter(annotations==TRUE & language == "english")
```

## Downloading documents

(Bulk-)Downloading documents from the Manifesto Corpus works via the function `mp_corpus(...)`.
This function understands different inputs. 

1. You can download election programmes on an individual basis
by listing combinations of party ids and election dates in a `data.frame`
and passing it to `mp_corpus(...)`:

```{r warning=FALSE}
wanted <- data.frame(party=c(41220, 41320),
                     date=c(200909, 200909))
mp_corpus(wanted)
```

The party ids (41220 and 41320 in the example) are the ids as in the Manifesto Project's
main dataset.
They can be found in the current dataset documentation at
[https://manifesto-project.wzb.eu/datasets]() or in the main dataset.

Note that we received only 1 document, while querying for two.
This is because the party with the id 41220 (KPD) did not run for elections
in September 2009.

2. Instead of typing all these combinations by hand, one can also do this easier: `mp_availability` returns a data frame in the same format as the `wanted` data frame above that is used to query the corpus. So, to get a corpus with all english-language annotated documents, you could just pass the object `english_annotated` saved above to the `mp_corpus` function:

```{r}
mp_corpus(english_annotated)
```

3. `mp_corpus` can be called with a logical expression specifying the subset of the Manifesto
Corpus that you want to download:

```{r, warning=FALSE}
my_corpus <- mp_corpus(countryname == "Austria" & date > 200100)
my_corpus
```

This queries for all documents in the Manifesto Corpus for Austrian parties published after or in the the year 2001. (The format of the date variable in the Manifesto Project Dataset is YYYYMM).

The variable names in the logical expression used for querying the corpus
database (`countryname` and `date` in the example above) can be any column names from the
Manifesto Project's Main Dataset or your current R environment. The Main Dataset
itself is available in manifestoR via the funcion. The following command queries all documents with [a "rile" score higher than 60](https://manifesto-project.wzb.eu/tutorials/main-dataset):

```{r, warning = FALSE}
mp_corpus(rile > 60) ## another example of data set based corpus query
```

A convenient way to download the whole Manifesto Corpus is to type `mp_corpus(TRUE)`. However, please have in mind that this might take a while to process and download. In most cases, it is sufficient to download a subset of the Manifesto Corpus.

## The ManifestoCorpus object

`mp_corpus` returns a `ManifestoCorpus` object, a subclass of `Corpus` as defined
in the natural language processing package `tm` (Feinerer & Hornik 2015).
Following `tm`s logic, a `ManifestoCorpus` consists of `ManifestoDocuments`.
Documents in corpus can be indexed via their `manifesto_id` (consisting of
the CMP party code, an underscore, and either the election year, if unambigous,
or the election year and month) or via their position in the corpus.
For both, corpus and documents, `tm` provides accessor functions to the corpus
and documents content and metadata:

```{r}
head(content(my_corpus[["42110_200211"]]))
head(content(my_corpus[[1]]))
meta(my_corpus[["42110_200211"]])
```




# Processing and analysing the corpus documents

As in `tm`, the textual content of a document is returned by the function `content`. In the documents with annotations, the text is stored as a character vector where each element is one quasi-sentence. For example the extract below are the first four quasi-sentences from the Austrian Green Party's manifesto from the 2006 election.

```{r}
txt <- content(my_corpus[["42110_200610"]])
class(txt)
head(txt, n = 4)
```

In documents with no annotations, the text is stored as a single very long character string. The following shows an extract of the German SPD manifesto from 1994 election.

```{r}
txt <- content(mp_corpus(party == 41320 & date == 199410)[["41320_199410"]])
substr(txt,19900,20512)
```

## Working with the CMP codings

The central way for accessing the CMP codings is the accessor method `codes(...)`.
It can be called on `ManifestoDocument`s and `ManifestoCorpus`s and returns a vector
of the CMP codings attached to the quasi-sentences of the document/corpus in a row:

```{r}
doc <- my_corpus[["42110_200610"]]
head(codes(doc), n = 15)
head(codes(my_corpus), n = 15)
```

Thus you can for example use R's functionality to count the codes or select quasi-
sentences (units of texts) based on their code:
```{r}
table(codes(doc))

doc_subcodes <- subset(doc, codes(doc) %in% c(202, 503, 607))
length(doc_subcodes)
length(doc_subcodes)/length(doc)
```

The CMP coding scheme can be found in the online documentation of the Manifesto
Project dataset at [https://manifesto-project.wzb.eu/coding_schemes/1](). Obviously, `codes()` only works on documents that are digitally annotated (annotations==TRUE).


## Using the document metadata

Each document in the Manifesto Corpus has meta information about itself attached.
They can be accessed via the function `meta`:
```{r}
meta(doc)
```

It is possible to access and also modify specific metadata entries:
```{r}
meta(doc, "party")
meta(doc, "manual_edits") <- TRUE
meta(doc)
```

Document metadata can also be bulk-downloaded with the function `mp_metadata`,
taking the same set of parameters as `mp_corpus`:
```{r warning=FALSE}
metas <- mp_metadata(countryname == "Spain")
head(metas)
```

The field ...

* ... `party` contains the party id from the Manifesto Project Dataset.
* ... `date` contains the month of the election in the same format as in the
Manifesto Project Dataset (`YYYYMM`)
* ... `language` specifies the language of the document as a word.
* ... `is_primary_doc` is FALSE only in cases where for a single party and
election date multiple manifestos are available and this is the document not used
for coding by the Manifesto Project.
* ... `may_contradict_core_dataset` is TRUE for documents where the CMP codings
in the corpus documents might be inconsistent with the coding aggregates in the
Manifesto Project's Main Dataset. This applies to manifestos which have been either
recoded after they entered the dataset or cases where the dataset entries are
derived from hand-written coding sheets used prior to the digitalization of the
Manifesto Project's data workflow, but the documents were digitalized and added
to the Manifesto Corpus afterwards.
* ... `annotations` is TRUE whenenver there are CMP codings available for the document.
* ... `has_eu_code` marks document in which the additional coding layer `eu_code` is present. These codes have been assigned to quasi-sentences by CMP coders additionally to the main CMP code to indicate policy statements that should or should not be implemented on the level of the European union.
* ... `handbook` indicates the version of the coding instructions that was used to annotate the document. See [this website](https://manifesto-project.wzb.eu/information/documents/handbooks) for more information on the (different versions of) coding instructions.
* ... `is_copy_of` In a few cases, we copy manifestos to use them for more than one party-date combination (eg. in case of some alliances). In such cases, this field indicates the manifesto-id of the original document. When doing computerized text analysis, you might often want to exclude these cases otherwise you will deal with duplicate documents. 
* ... `title` is the title of the document.

The other metadata entries have primarily technical functions for communication
between the `manifestoR` package and the online database.


## Working with additional layers of codings

Besides the main layer of CMP codings, you can create, store and access additional
layers of codings in `ManifestoDocument`s by passing a name of the coding layer 
as additional argument to the function `codes()`:

```{r}
## assigning a dummy code of alternating As and Bs
codes(doc, "my_code") <- rep_len(c("A", "B"), length.out = length(doc))
head(codes(doc, "my_code"))
```

You can view the names of the coding layers stored in a `ManifestoDocument` with
the function `code_layers()`:
```{r}
code_layers(doc)
```

Note that certain documents downloaded from the Manifesto Corpus Database already
have a second layer of codes named `eu_code`. These are codes that have been assigned
to quasi-sentences by CMP coders additionally to the main CMP code to indicate
policy statements that should or should not be implemented on the level of the
European union. The documents that were coded in this way are marked in the
corpus' metadata with the flag `has_eu_code`
(see below [Using the document metadata](#using-the-document-metadata)).
Note that, since these codes also have been used for computing the `per` and `rile`
variables in the Manifesto Project Main Dataset, they are also used in `manifestoR`s
`count_codes` and `rile` functions (see below [Scaling texts](#scaling-texts)) if the respective metadata flag is present.

## Text mining tools

Since the Manifesto Corpus uses the infrastructure of the `tm` package
(Feinerer & Hornik 2015), all of `tm`s filtering and transformation functionality
can be applied directly to the downloaded `ManifestoCorpus`.

For example, standard natural language processors are available to clean the corpus:

```{r}
head(content(my_corpus[["42110_200809"]]))
corpus_cleaned <- tm_map(my_corpus, removePunctuation)
corpus_nostop <- tm_map(corpus_cleaned, removeWords, stopwords("german"))
head(content(corpus_nostop[["42110_200809"]]))
```

So is analysis in form of term document matrices:
```{r}
tdm <- TermDocumentMatrix(corpus_nostop)
inspect(tdm[c("menschen", "wahl", "familie"),])
findAssocs(tdm, "stadt", 0.97) ## find correlated terms, see ?tm::findAssocs
```

For more information about the functionality provided by the `tm`,
please refer to its [documentation](http://cran.r-project.org/web/packages/tm/vignettes/tm.pdf).

## Selecting relevant parts of text

For applications in which not the entire text of a document is of interest, but
rather a subset of the quasi-sentences matching certain criteria,
`manifestoR` provides a function `subset(...)` working just like R's internal
`subset` function.

It can, for example, be used to filter quasi-sentences based on codes or the text:

```{r}
# subsetting based on codes (as example above)
doc_subcodes <- subset(doc, codes(doc) %in% c(202, 503, 607))
length(doc_subcodes)

# subsetting based on text
doc_subtext <- subset(doc, grepl("Demokratie", content(doc)))
head(content(doc_subtext), n = 3)
head(codes(doc_subtext), n = 10)
```

Via `tm_map` the filtering operations can also be applied to an entire corpus:
```{r}
corp_sub <- tm_map(my_corpus, function(doc) {
  subset(doc, codes(doc) %in% c(202, 503, 607))
})
head(content(corp_sub[[3]]))
head(codes(corp_sub))
```

For convenience, it is also possible to filter quasi-sentences with specific
codes directly when downloading a corpus. For this, the additional argument
`codefilter` with a list of CMP codes of interest is passed to `mp_corpus`:
```{r, warning=FALSE}
corp_sub <- mp_corpus(countryname == "Australia", codefilter = c(103, 104))
head(content(corp_sub[[1]]))
head(codes(corp_sub))
```



# Viewing original documents

Apart from the machine-readable, annotated documents, the Manifesto Corpus also
contains original layouted election programmes in PDF format. If available, they
can be viewed via the function `mp_view_originals(...)`, which takes exactly the
format of arguments as `mp_corpus(...)` ([see above](#downloading-documents)), e.g.:

```{r eval = FALSE}
mp_view_originals(party == 41320 & date == 200909)
```

The original documents are shown in you system's web browser. All URLs opened
by this function refer only to the Manifesto Project's Website. If you want to
open more than 5 PDF documents at once, you have to specify the maximum number
of URLs allows to be opened manually via the parameter `maxn`. Since opening
URLs in an external browser costs computing resources on your local machine,
make sure to use only values for `maxn` that do not slow down or make your computer
unresponsive.

```{r eval = FALSE}
mp_view_originals(party > 41000 & party < 41999, maxn = 20)
```


# Efficiency and reproducibility: caching and versioning

To save time and network traffic, `manifestoR` caches all downloaded data and
documents in your computer's working memory and connects to the online database
only when data is required that has not been downloaded before.

```{r eval=TRUE, include=FALSE}
mp_emptycache()
```

```{r message=TRUE, warning=FALSE}
corpus <- mp_corpus(wanted)
subcorpus <- mp_corpus(wanted[3:7,])
```

Note that in the second query no message informing about the connection to the
Manifesto Project's Database is printed, since no data is actually downloaded.

This mechanism also ensures **reproducibility** of your scripts, analyses
and results: executing your code again will yield the same results, even if
the Manifesto Project's Database is updated in the meantime.
Since the cache is only stored in the working memory, however, in order to ensure
reproducibility across R sessions, it is advisable to
**save the cache to the hard drive** at the end of analyses and load it in the
beginning:

```{r eval=FALSE}
mp_save_cache(file = "manifesto_cache.RData")

## ... start new R session ... then:

library(manifestoR)
mp_setapikey("manifesto_apikey.txt")
mp_load_cache(file = "manifesto_cache.RData")
```

This way `manifestoR` always works with the same snapshot of the Manifesto Project
Database and Corpus, saves a lot of unnecessary online traffic and also enables
you to continue with your analyses offline.

Each snapshot of the Manifesto Corpus is identified via a version number, which is
stored in the cache together with the data and can be accessed via
```{r}
mp_which_corpus_version()
```

When collaborating on a project with other researchers, it is advisable to use
the same corpus version for reproducibility of the results.
`manifestoR` can be set to use a specific version with the functions
```{r}
mp_use_corpus_version("2015-3")
```

In order to guarantee reproducibility of **published work**, please also mention
the corpus version id used for the reported analyses in the publication.

For updating locally cached data to the most recent version of the
Manifesto Project Corpus, `manifestoR` provides two functions:

```{r}
mp_check_for_corpus_update()
mp_update_cache()
mp_check_for_corpus_update()
```

For more detailed information on the caching mechanism and on how to use and load
specific snapshots of the Manifesto Corpus, refer to the R documentations of the
functions mentioned here as well `mp_use_corpus_version`, `mp_corpusversions`,
`mp_which_corpus_version`.


# Exporting documents

If required `ManifestoCorpus` as well as `ManifestoDocument` objects can be
converted to R's internal `data.frame` format and processed further:
```{r}
doc_df <- as.data.frame(doc)
head(within(doc_df, {
  ## for pretty printing
  text <- paste0(substr(text, 1, 60), "...")
}))
```

The function also provides a parameter to include all available metadata in
the export:
```{r}
doc_df_with_meta <- as.data.frame(doc, with.meta = TRUE)
print(names(doc_df_with_meta))
```

Note again that also all functionality provided by `tm`, such as `writeCorpus`
is available on a `ManifestoCorpus`.


# Additional Information

When publishing work using the Manifesto Corpus, please make sure to cite it correctly and to give the identification number of the corpus version used for your analysis. You can print citation and version information with the function `mp_cite()`.

For a more detailed reference and complete list of the functions provided
by `manifestoR`, see the R package reference manual on CRAN:
[http://cran.r-project.org/web/packages/manifestoR/manifestoR.pdf](http://cran.r-project.org/web/packages/manifestoR/manifestoR.pdf)


\newpage

# References

Feinerer, I., & Hornik, K. (2015). Tm: Text Mining Package. http://cran.r-project.org/web/packages/tm/index.html
